---
title: "wk05_ggplot_Assigment"
author: "Alex Smith"
date: "February 5, 2016"
output: 
  html_document: 
    toc: yes
---

##Joining Data

```{r}


if ('env-info_hw' %in% list.files() & interactive()){
    setwd('env-info_hw' )
}
```

per capita CO2 emissions
You’ll join the population dataset to calculate per capita CO2 emissions.

**Task.** Calculate the per capita emissions by country (not World or EU28) and return the top 5 emitting countries for 2014.
```{r}

library(readr)
library(readxl) # install.packages('readxl')
library(tidyr)
library(httr)   # install.packages('httr')
library(plyr)   # install.packages('plyr')
library(dplyr)

print(getwd())

# population Excel file
url = 'http://edgar.jrc.ec.europa.eu/news_docs/CO2_1970-2014_dataset_of_CO2_report_2015.xls'
out = 'co2_europa.xls'
if (!file.exists(out)) writeBin(content(GET(url), 'raw'), out)


# country lookup csv file
url = 'https://raw.githubusercontent.com/ucsb-bren/env-info/gh-pages/data/co2_country_to_popn.csv'
out = 'co2_country_to_popn.csv'
if (!file.exists(out)) writeBin(content(GET(url), 'raw'), out)

xls='../data/co2_europa.xls'

# read in carbon dioxide emissions file
co2 = read_excel(xls, skip=12) %>%
   gather("year","emiss", 2:46) %>%
    mutate(year=as.numeric(year))
     


# get lookup table to translate some country names in popn to those in co2 table
cntry = read_csv('co2_country_to_popn.csv')

# read in population and revalue some country names to match co2 table
popn = read_csv('https://raw.githubusercontent.com/datasets/population/master/data/population.csv') %>%
  select(Country=`Country Name`, year=Year, popn=Value) %>%
  mutate(
    Country = plyr::revalue(
      Country, 
      setNames(cntry$Country_co2, cntry$Country_popn)))
# rename(popn,country=Country)

percap1=inner_join(co2, popn, by=c("Country", "year")) %>%
  filter(year==2014) %>%
    filter(!Country==c("World", "EU28")) %>%
      mutate(PCemiss= emiss/popn) %>%
        arrange(desc(PCemiss)) %>%
          slice(1:5) %>%
            print()
        
```

  
  
**Task.** Summarize the per capita emissions by country (not World or EU28) as the mean (ie average) value across all years and return the top 5 emitting countries.


```{r}

percap2=inner_join(co2, popn, by=c("Country", "year")) %>%
    filter(!Country==c("World", "EU28")) %>%
      mutate(PCemiss= emiss/popn) %>%    
        group_by(Country) %>%    
          summarise(
            Mean=mean(PCemiss)
            )    %>%
               arrange(desc(Mean)) %>%      
                slice(1:5) %>%
            print()
```














##Plots

```{r dataframes, include=FALSE}
library(dplyr)
library(lubridate)
library(ggplot2)
library(stats)
library(readr)

# thin=read.csv('../data/thinning.csv') %>%
#   mutate(monthpost=paste( thin$postyears,  thin$month, sep="."))

thin = read_csv('../data/thinning.csv') %>%
  mutate(postfracyears = postyears + month/12)
  
clim=read.csv('../data/clim.csv')
```


```{r plot_1, echo=FALSE}

# p1=ggplot(thin, aes(x=as.numeric(monthpost), y=cpool,     col=(as.factor(postyears)))) +
#    stat_summary(fun.y="mean", geom = "line")

thin %>%
  ggplot(aes(
    x   = month, 
    y   = cpool,
    color = as.factor(postyears))) +
  stat_summary(fun.y = "mean", geom = "line") +
  stat_summary(mapping = aes(
    x   = month, 
    y   = cpool), fun.y="mean", 
    geom = "line", color='black', size=2) +
  labs(x="month", y="carbon sequesterd")+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

## Group Data

see [group page](https://fishmgt.github.io/wk_05_ggplot_AS.html)