---
title: "wk_06_widgets"
author: "Alex Smith"
date: "February 12, 2016"
output: 
  html_document: 
    toc: yes
---


##Group Data: Regional Catch Data Motion Chart (1990-2013)

    
####See more fun @fishmgt interactive plots [here](http://fishmgt.github.io/wk_06_widgets_JHC.html)

```{r googlevis library, results='asis', warning=F, include=F}
suppressPackageStartupMessages({
  library(googleVis)  # install.packages('googleVis')
})
library(readr)
library(readxl)
library(ggplot2)
library(plyr)
library(dplyr)
library(tidyr)

# must set this option for googleVis charts to show up
op <- options(gvis.plot.tag='chart')

```


```{r motion data only, include=F}

motioncatch=read_csv('../../fishmgt.github.io/data/For CRSeafood code/motioncatch.csv') %>%
  distinct() %>%
  group_by(Region,Year) %>%
  summarise(
    
    Annual_Catch=sum(Catch)
  )

```

```{r MotionChartCatch Data, results='asis', tidy=FALSE, echo=F}

C= gvisMotionChart(motioncatch, idvar='Region', timevar = 'Year',  sizevar = 'Annual_Catch',  options=list(state='{"playDuration":15000,"xLambda":1,"yAxisOption":"2","iconType":"BUBBLE","xAxisOption":"_TIME","xZoomedDataMax":1356998400000,"yZoomedDataMin":506,"orderedByX":false,"yZoomedIn":false,"sizeOption":"2","yLambda":0,"uniColorForNonSelected":false,"showTrails":true,"iconKeySettings":[],"xZoomedDataMin":631152000000,"orderedByY":false,"nonSelectedAlpha":0.4,"duration":{"multiplier":1,"timeUnit":"Y"},"colorOption":"_UNIQUE_COLOR","dimensions":{"iconDimensions":["dim0"]},"time":"1990","xZoomedIn":false,"yZoomedDataMax":634463};'))


plot(C)

```

##Individual Tasks  

###**Task 1**

What is the effect of changing the `adjust` parameter on the line density?

We can use the `manipulate` function to provide interactive sliders, checkboxes and pickers. It only works within RStudio and does not work in a knitting context, so be sure to set `eval=FALSE` in the R chunk options.



```{r manipulate_before, eval=F}
library(manipulate) # install.packages('manipulate')

manipulate({
  
  faithful %>%
    ggplot(aes(eruptions)) + 
    geom_histogram(aes(y = ..density..),  bins = 20) +
    geom_density(color='blue', size=2, adjust=a) +
    xlab('duration (minutes)') +
    ylab('frequency density') +
    ggtitle('geyser eruption duration')
  
}, a = slider(min = 0, max = 2, initial = 1, label = 'bandwidth adjustment', step = 0.2))
```

 **Task.** Add another R chunk with a slider adjusting the number of bins from 5 to 50, with step increments of 5.

```{r manipulat_after, eval=F}

manipulate({
  
  faithful %>%
    ggplot(aes(eruptions)) + 
    geom_histogram(aes(y = ..density..),  bins = b) +
    geom_density(color='blue', size=2, adjust=a) +
    xlab('duration (minutes)') +
    ylab('frequency density') +
    ggtitle('geyser eruption duration')
  
}, a = slider(min = 0, max = 2, initial = 1, label = 'bandwidth adjustment', step = 0.2), 
   b = slider(min= 5, max= 50, initial=20, label = 'bin width adjustment', step= 5))
```



###**Task 2**


Let's use ggvis `tooltip()` to show values of a scatterplot on mouse hover.

```{r ggvis add_tooltip_before, eval=FALSE}
library(ggvis) # install.packages('ggvis')

cars = mtcars %>%
  add_rownames('model') %>%        # dplyr drops rownames
  mutate(id = row_number()) # add an id column to use ask the key

all_values <- function(x) {
  if(is.null(x)) return(NULL)
  row <- cars[cars$id == x$id, ]
  paste0(names(row), ": ", format(row), collapse = "<br/>")
}

cars %>% 
  ggvis(x = ~wt, y = ~mpg, key := ~id) %>%
  layer_points() %>%
  add_tooltip(all_values, 'hover')
```

**Task.** Add another R chunk that only applies the `add_tooltip` on mouse click.

```{r ggvis add_tooltip_after, eval=FALSE}

cars = mtcars %>%
  add_rownames('model') %>%        # dplyr drops rownames
  mutate(id = row_number()) # add an id column to use ask the key

all_values <- function(x) {
  if(is.null(x)) return(NULL)
  row <- cars[cars$id == x$id, ]
  paste0(names(row), ": ", format(row), collapse = "<br/>")
}

cars %>% 
  ggvis(x = ~wt, y = ~mpg, key := ~id) %>%
  layer_points() %>%
  add_tooltip(all_values, 'click')

```


###**Task 3**

 `datatable`

But there are many more rows than just the first 6. Which is why an interactive widget could be so helpful.

```{r datatable_before}
library(DT) # install.packages('DT')

# default datatable
datatable(iris)

# remove document elements
datatable(iris, options = list(dom = 'tp'))
```

Note how the [`dom` option](http://datatables.net/reference/option/dom) removed other elements from display such as the Show (`l`ength), Search (`f`iltering) and Showing (`i`nformation) elements from the default, but kept the `t`able and `p`agination control.

**Task.** Output the table again with `datatable` and set the `options` to have `pagelength` of just 5 rows. (See `?datatable` and <http://rstudio.github.io/DT/>).

```{r datatable_after}


# remove document elements
datatable(iris, options = list(pageLength = 5, dom = 'tip'))
```